---
title: "World Mean (FGDL, IPSL, MPIL) MCP for RCP 2.6"
author: "Juliano Palacios Abrantes"
date: "04/05/2018"
output: html_document
---

# Overall approach
I first average the DBEM's Maximum Catch Potential (MCP) of 2005 to 2014 (last years of fisheries data from SAU) for each species for each one of the circulation models that we work with (GFDL, MPIL, IPSL). Secondly, I average the 2005-2014 average for all models. Finally, I added the result of each specie within each grid-cell of the world map. 

The results represent the mean aggregated MCP of all species in our database for each grid-cell of the world.


```{r setup, eval=T, echo=F, warning=F,message=F}

#### READ ME !!! ####
# Run this chunk before knit so you make sure you have all pkgs installed in R
# This analysis will only run on a computer connected to the Drobo. 
# If you want to run it you will need to change the "path" of the DBEM_imp_F.R function


#### Set up packages 

ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}


#### Library ####
packages <- c(
  "readxl", # Read dataframe
  "dplyr", # Data manipulation
  "tidyr", # Data manipulation
  "ggplot2", #Nice grpahs and spatial analysis
  # "rgdal",
  # "RColorBrewer", # For more nice collors
  "knitr",
  "kableExtra", # For Nice tables
  "data.table",
  # "ggrepel",
  # "gridExtra",
  # "ggmap",
  "rgdal", #Spatial analysis 
  "tools" #Spatial analysis 
)

ipak(packages)

```

```{r Data_Functions_Loading, eval = T, echo = T, warning=F,message=F}

# Data needed

#List of DBEM species
exploited_species <- fread("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv") 

# Coordenates for DBEM mapping
Coor <- fread("~/Documents/UBC/Oceans_Project/Distribution/DBEM/Data/Lat_Lon_DBEM.txt",header = FALSE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")

# Functions
source("./Functions/DBEM_imp_F.R") #Imports a single species

```


```{r DBEM_Data, eval=T, echo =F}
# IT works untill now for 1 model

# For each species...
# Step 1: Average the MCP of 2005 to 2014
# Step 2: Mean of all three models (GFDL, IPSL, MPI)
# Step 3: Addition of averages in each INDEX


#### Parameters ####


for(r in 2:2){
  
  if(r == 1){
    Models = c("GFDL26F1","IPSL26F1","MPI26F1") #Low RCP
    Name = paste("World_MCP_05-14_26.csv")
  }else{
    Models = c("GFDL85F1","IPSL85F1","MPI85F1") #High RCP
    Name = paste("World_MCP_05-14_85.csv")
  }
  
  # Dummy final dataset
  Final_Dataset <- Coor %>% 
    select(INDEX) %>% 
    mutate(Total = 0)
  
  # Analysis ####
  for(s in 1:nrow(exploited_species)){ # For all species
    
    NoSpeciesError <-  tryCatch({ # <- This will solve for species not modeled in the overall list
      
      for(m in 1:length(Models)){ # For all models
        
       # Step 1. Average MCP of 2046 - 2055 ####
        # 1.1 Imports DBEM data
        
        x <-dbem_Import(exploited_species[s,1],
                        2005,
                        2014,
                        Model = Models[m],
                        Data_Type= "Catch"
        )
        
        # 1.2 Averages 10 years of data
        Mean_Spp <- x %>% 
          mutate(Mean = rowMeans(.[,2:11], na.rm = T)) %>% 
          select(INDEX,
                 Mean)
        
        Mean_Spp <- as.data.table(Mean_Spp) # Set as datatable 
        
        # 1.3 Builds means for each model in one dataset
        if(m == 1){
          Mean_Data <- copy(Mean_Spp) # <- copies the previouse data.table
          
        }else{
          setkey(Mean_Data, # <- sets the data.table as "reference" ?setkey
                 INDEX); setkey(Mean_Spp,
                                INDEX);
          list <- list(Mean_Data, # <- creates a list to merge the tables
                       Mean_Spp)
          
          Mean_Data <-merge(Mean_Data, #<- Merges all the data in one single file.
                            Mean_Spp,
                            all = TRUE) # <- This allows for the NA's
        }
        
      } # Close Models loop end
      
      #### CHECKPOINT- CHARLY- Step 1 ####
      # Run it from the "Model loop" to see if the loop is averaging all three models result
      # }
      # Mean_Data should be a dataset with 4 columns: INDEX and one with the 10y mean of each model (check)
      # Manually checked that each model was in fact averaged for each species using 600243
      ###_________________ END OF CHECKPOINT _______________ ###
      
      
      # Step 2. Average MCP of all models ####
      Average <- Coor %>% # To do it by INDEX
        left_join(Mean_Data,
                  by ="INDEX") %>% 
        mutate(Model_Mean = rowMeans(.[,4:6], na.rm = T)) %>% # Averages the values (1-3 Index, lat, long)
        mutate(Final_Mean = ifelse(is.na(Model_Mean),0,Model_Mean)) %>%  # Converts NA's to 0 for future sum
        select(INDEX,
               Final_Mean)
      
      
      #### CHECKPOINT- CHARLY- Step 2 ####
      # Average should be a dataset for the world with the mean of all models for each grid (Check)
      # Check that the NA's are ceros and numeric for futher addition (Check)
      ###_________________ END OF CHECKPOINT _______________ ###
      
      
      # Step 3. Adding the Averages ####
      
      Final_Dataset[,2] = Final_Dataset[,2] + Average[,2]
      
     # End of NoSpeciesError TryCatch function
    },
    error=function(e){
      e
      print(paste("No information for",exploited_species[s,1],Models[m],sep = " "))
    }
    )
    
    #### CHECKPOINT- CHARLY- Step 3 and NoSpeciesError ####
    # Final_Dataset should be the addition of each species 10 years mean of all models
    # Manually checked that each model was in fact added for each gridcell using 600243 and 600244
    # Cells that have value for each species are added as well as cells that ahve only one value (for one spp) 
    
    # Error check should jump those species in the list that do not have data. (Cheked)
    # Tested with 600003 (no data), 600243 (data) and 600244 (overlapping data) 
    ###_________________ END OF CHECKPOINT _______________ ###
    
    #### Miscellaneous ####
    
    # Just to have a sense of time...
    Timer <- paste(s,"Species Completed, ",nrow(exploited_species)-s,"Species Left")
    print(Timer)
    print(s+1)
  }
  
  # Last step Add the coordinates
  
  Final_Dataset <- Final_Dataset %>% 
    left_join(Coor,
              by = "INDEX")
  
  
  write.csv(Final_Dataset,
            Name, #World MCP between 2004 and 2014 for Low emission RCP (2.6)
            row.names = FALSE)
}


```

# Plotting the Map

```{r Maps_Data, eval = T, echo = F, warning=F,message=F}

# The path
path_world <- ("/Users/jpalacios/Documents/R CODE/Spatial_Analysis/Data/ne_10m_coastline")
#The File
fnam_world <- "ne_10m_coastline.shp"

#Load it!
World <- readOGR(dsn = path_world,
                 layer =file_path_sans_ext(fnam_world))

# Functions needed

# Transforms shapefile to dataframe
fortify.shape <- function(x){
  x@data$id <- rownames(x@data)
  x.f = fortify(x, region = "id")
  x.join <- inner_join(x.f, x@data, by = "id")
}

```

```{R The_map, eval =T, echo = T, warning=F,message=F}


#### Option A ####

Forty_World<- fortify(World)

Map <- ggplot() +
  geom_path(data = Forty_World, aes(x = long,
                                    y = lat,
                                    group = group
  ),
  color = "black",
  size = 0.25) +
  coord_map(projection = "mercator") +
  labs(list(title = "",
            x = "Longitude",
            y = "Latitude")) +
  theme_classic() +
  geom_tile(data = filter(Final_Dataset, Total >1), # And now the DBEM mean data
            aes(x = Longitude,
                y = Latitude,
                fill = Total,
                colour = Total
            )
  )

ggsave("Map_8.5.png",
       plot = Map,
       width = 25,
       height = 10,
       units = "in"
)

```

```{R Double_Check_Data, eval =F, echo = T, warning=F,message=F}

Low_Data <- read.csv("~/Documents/Dropbox/Global_Andres/World_MCP_04-14_26.csv") %>% 
  filter(Total > 0)
High_Data <- read.csv("~/Documents/Dropbox/Global_Andres/World_MCP_04-14_85.csv") %>% 
  filter(Total > 0)

# Select Data to test

Test_Data <- High_Data

# Basic stats

Max_L <- max(Test_Data$Total) # 7532.24
Min_L <- min(Test_Data$Total) # 4.12-e06
Mean_L <- mean(Test_Data$Total)
SD_L <- sd(Test_Data$Total)

# Hist

ggplot(Test_Data) +
  geom_histogram(
    aes(
      x= Total
    ),
    binwidth = 50)

# Ok, veeeery squeewed


# First Test. Selecting only extreemly big values

PlotData = filter(Test_Data, Total >1000)
max(PlotData$Total)

# All the values over 1000 are in the Humboldt current

# Second. Removing extreemly big values
PlotData = filter(Test_Data, Total <1000)
max(PlotData$Total)

# Third. Maybe using log values

PlotData <- Test_Data %>% 
  mutate(Log = log10(Total))

ggplot() +
  geom_path(data = Forty_World, aes(x = long,
                                    y = lat,
                                    group = group
  ),
  color = "black",
  size = 0.25) +
  coord_map(projection = "mercator") +
  labs(list(title = "",
            x = "Longitude",
            y = "Latitude")) +
  theme_classic() +
  geom_tile(data = PlotData, # And now the DBEM mean data
            aes(x = Longitude,
                y = Latitude,
                fill = Log,
                colour = Log
            )
  ) +
  coord_map(projection = "mercator") +
  scale_colour_gradient2(
    limits=c(-4,
             2),
    breaks = 0.5) +
  scale_fill_gradient2(
    limits=c(-4,
             2)
  )


ggsave("Map_Test_I85.png",
       plot = last_plot(),
       width = 25,
       height = 10,
       units = "in"
)

```

